### 2014-03-18 manuscript.tex

In this Section we present two methods to classify sentences to either positive, neutral or negative towards a keyword. Both methods follow the same general algorithm presented below which is inspired by Ding et al. \shortcite{Ding08} and is based on a sentiment lexicon. Both keywords, sentiment words and sentiment shifters can in general appear several times in a sentence. Sentiment shifters is words that potentially shift the sentiment of a sentence from positive to negative or negative to positive. E.g. ``not happy'' have the opposite polarity than just ``happy''. We assume that we have a list of sentiment words each associated with a sentiment score representing the polarity and strength of the sentiment word. Let $kw_i, i \in \{1,2,\ldots,I\}$ represent appearance number $i$ of the keyword in the sentence. Further let $sw_{jk}, j \in \{1,2,\ldots,J\}, k \in\{1,2,\ldots,K_j\}$ be appearance number $k$ of sentiment word $j$. Thus, sentiment word $j$ appears a total of $K_j$ times in the sentence. Finally let $ss_{lm}, l \in \{1,2,\ldots,L\}, m \in\{1,2,\ldots,M_l\}$ be appearance number $m$ of sentiment shifter $l$. We 
compute a sentiment score, $S$, for the sentence as follows
\begin{equation}
  \label{eq:1}
  S = \frac{1}{I}\sum_{i=1}^{I} \sum_{j=1}^{J} \sum_{k=1}^{K}\mathbf{imp}(kw_i, sw_{jk})\mathbf{s}(sw_{jk}, \mathbf{ss})    
\end{equation}
where the function Imp computes the importance of the sentiment word $sw_{jk}$ on the keyword appearance $kw_i$. This will be computed in different ways as described below. Further, the function $\mathbf{s}$ computes whether the sentiment of $sw_{jk}$ should be shifted based on all the sentiment shifters in the sentence, denoted by $\mathbf{ss}$. The function $\mathbf{s}$ returns $-1$ (sentiment shift) if some of the sentiment shifters is within $d_{p}$ words in front or $d_{n}$ words behind $sw_{jk}$, respectively. Else $\mathbf{s}$, returns $1$. We classify the opinion towards the keyword to be positive, neutral or negative if $S >= t_p$,  $t_p > S > t_n$ and  $S <= t_n$, respectively. The parameters $d_p, d_n, t_p$ and $t_n$ is tuned using a training set.

@@

For the word distance method we use the following Imp function
\begin{equation}
  \label{eq:2}
  \mathbf{imp}(kw_i, sw_{jk}) = \frac{\mathbf{sentsc}(sw_{jk})}{\mathbf{worddist}(kw_i, sw_{jk})}
\end{equation}
where $\mathbf{sentsc}(sw_{jk})$ is the sentiment score of $sw_{jk}$ from the sentiment lexicon and $\mathbf{worddist}(kw_i, sw_{jk})$ is the number of words between $kw_i$ and $sw_{jk}$ in the sentence plus one.

@@

Let $\mathcal{D}$ denote the set of all important grammatical relations. The function $\mathbf{gram}(kw_i, sw_{jk})$ returns the grammatical relation, and if $\mathbf{gram}(kw_i, sw_{jk}) \in \mathcal{D}$, then the function $W_{\mathbf{dep}}(kw_i, sw_{jk}) \in [0,1]$, return the importance of the grammatical relation. Further let $\mathbf{treedist}(kw_i, sw_{jk})$ return the number of words between the two words in the parse tree plus one. The Imp function is computed as follows. If $\mathbf{gram}(kw_i, sw_{jk}) \in \mathcal{D}$ we set 
\begin{align}
  \begin{split}
  \mathbf{imp}&(kw_i, sw_{jk}) = \\
  &\alpha \cdot \mathbf{sentsc}(sw_{jk}) W_{\mathbf{dep}}(kw_i, sw_{jk}) \\
  +&(1 - \alpha) \cdot \dfrac{\mathbf{sentsc}(sw_{jk})}{\mathbf{treedist}(kw_i, sw_{jk})}    
  \end{split}
\end{align}
where $\alpha \in [0,1]$ is parameter that weights the score from the important dependence path and the tree distance. If $\mathbf{gram}(kw_i, sw_{jk}) \not\in \mathcal{D}$ we simply set 
\begin{align}
  \begin{split}
  \mathbf{imp}(kw_i, sw_{jk}) = \dfrac{\mathbf{sentsc}(sw_{jk})}{\mathbf{treedist}(kw_i, sw_{jk})}    
  \end{split}
\end{align}

@@

We automatically split the material into sentences and all sentences containing some of the keywords were selected for further analysis. If a sentence contained more than one keyword, e.g. both islam and muslim, the sentence were repeated one time for each key word. We could then measure the opinion towards each of the keywords in the sentence separately.

Each sentence were manually annotated whether the commenter were positive, negative or neutral towards the key word in the sentence. Each sentence were evaluated individually. The sentences were annotated based on all our knowledge, e.g. a sentence like ``Muhammed is like Hitler'' would be annotated as a negative opinion towards Muhammed. Further, if a commenter presented a negative fact about the keyword, the sentence would be denoted as negative.\\
INTERANNOTARY!!\\
Finally the material were divided in to two parts where the first half of the debate articles with subsequent discussions where in the training set and the rest were the test set. The individuals working on finding the important dependence paths, did only use the training set and did never see the test set before the decisions about dependence paths were decided.

\subsection{Sentiment lexicon}

% Resultater

@@

>>>>

In this Section we present two methods to classify sentences to either positive, neutral or negative towards a keyword. Both methods follow the same general algorithm presented below which is inspired by Ding et al. \shortcite{Ding08} and is based on a sentiment lexicon. Both keywords, sentiment words and sentiment shifters can in general appear several times in a sentence. Sentiment shifters is words that potentially shift the sentiment of a sentence from positive to negative or negative to positive. E.g. ``not happy'' have the opposite polarity than just ``happy''. We assume that we have a list of sentiment words each associated with a sentiment score representing the polarity and strength of the sentiment word (sentiment lexicon). Let $kw_i, i \in \{1,2,\ldots,I\}$ represent appearance number $i$ of the keyword in the sentence. Further let $sw_{j}, j \in \{1,2,\ldots,J\}$ be appearance number $j$ of a sentiment word in the sentence. Finally let $\textbf{ss} = (ss_1, ss_2, \ldots, ss_K)$ repressent the sentiment shifters in the sentence. We compute a sentiment score, $S$, for the sentence as follows
\begin{equation}
  \label{eq:1}
  S = \frac{1}{I}\sum_{i=1}^{I} \sum_{j=1}^{J} \mathbf{imp}(kw_i, sw_{j})\mathbf{shift}(sw_{j}, \mathbf{ss})    
\end{equation}
where the function Imp computes the importance of the sentiment word $sw_{j}$ on the keyword appearance $kw_i$. This will be computed in different ways as described below. Further, the function $\mathbf{shift}(sw_{j}, \mathbf{ss})$ computes whether the sentiment of $sw_{j}$ should be shifted based on all the sentiment shifters in the sentence. It returns $-1$ (sentiment shift) if some of the sentiment shifters is within $d_{p}$ words in front or $d_{n}$ words behind $sw_{j}$, respectively. Else the function, returns $1$ (no sentiment shift). We classify the opinion towards the keyword to be positive, neutral or negative if $S >= t_p$,  $t_p > S > t_n$ and  $S <= t_n$, respectively. The parameters $d_p, d_n, t_p$ and $t_n$ is tuned using a training set.

@@

For the word distance method we use the following Imp function
\begin{equation}
  \label{eq:2}
  \mathbf{imp}(kw_i, sw_{j}) = \frac{\mathbf{sentsc}(sw_{j})}{\mathbf{worddist}(kw_i, sw_{j})}
\end{equation}
where $\mathbf{sentsc}(sw_{j})$ is the sentiment score of $sw_{j}$ from the sentiment lexicon and $\mathbf{worddist}(kw_i, sw_{j})$ is the number of words between $kw_i$ and $sw_{j}$ in the sentence plus one.

@@

Let $\mathcal{D}$ denote the set of all important grammatical relations. The function $\mathbf{gram}(kw_i, sw_{j})$ returns the grammatical relation, and if $\mathbf{gram}(kw_i, sw_{j}) \in \mathcal{D}$, then the function $W_{\mathbf{dep}}(kw_i, sw_{j}) \in [0,1]$, return the importance of the grammatical relation. Further let $\mathbf{treedist}(kw_i, sw_{j})$ return the number of words between the two words in the parse tree plus one. The Imp function is computed as follows. If $\mathbf{gram}(kw_i, sw_{j}) \in \mathcal{D}$ we use
\begin{align}
  \begin{split}
    \label{eq:3}
  \mathbf{imp}&(kw_i, sw_{j}) = \\
  &\alpha \cdot \mathbf{sentsc}(sw_{j}) W_{\mathbf{dep}}(kw_i, sw_{j}) \\
  +&(1 - \alpha) \cdot \dfrac{\mathbf{sentsc}(sw_{j})}{\mathbf{treedist}(kw_i, sw_{j})}    
  \end{split}
\end{align}
where $\alpha \in [0,1]$ is parameter that weights the score from the important dependence path and the tree distance may be tuned from a training set. If $\mathbf{gram}(kw_i, sw_{j}) \not\in \mathcal{D}$ we simply use 
\begin{equation}
  \label{eq:4}
  \mathbf{imp}(kw_i, sw_{j}) = \dfrac{\mathbf{sentsc}(sw_{j})}{\mathbf{treedist}(kw_i, sw_{j})}    
\end{equation}
Note that when $\alpha = 0$, \eqref{eq:3} reduces to \eqref{eq:4}.

\subsection{Statistical analysis of classification performance}

We compare the classification performance of the methods above using random effect logistic regression. Let the stochastic variable $Y_{tm} \in \{0,1\}$ represents whether method $m \in 1,2,\ldots,M$ classified correct opinion to sentence number $t \in \{1,2,\ldots,T\}$, where $T$ is the number of sentences in the test set. We let $Y_{tm}, t = 1,2,\ldots,T, m = 1,2,\ldots,M$ being the 0-1 dependent variable of the regression model.

@@

We automatically divided the material into sentences and all sentences containing at least one keyword and one sentiment word were kept for further analysis. If a sentence contained more than one keyword, e.g. both islam and quran, the sentence were repeated one time for each keyword. We could then classify the opinion towards each of the keywords in the sentence consecutively.

Each sentence were manually annotated whether the commenter were positive, negative or neutral towards the keyword in the sentence. Each sentence were evaluated individually. The sentences were annotated based on all our knowledge, e.g. a sentence like ``Muhammed is like Hitler'' would be annotated as a negative opinion towards Muhammed. Further, if a commenter presented a negative fact about the keyword, the sentence would be denoted as negative.\\
INTERANNOTARY!!\\
Finally the material were divided in to two parts where the first half of the debate articles with subsequent discussions where in the training set and the rest were in the test set. The researcher working on finding the important dependence paths, did only use the training set and did never see the test set before the decisions about dependence paths were decided. After the division, the training and test set consisted of a total of 382 and 308 sentences, respectively. Table \ref{tab:1} summarizes the opinions in the sentences
\begin{table}
  \centering
  \begin{tabular}{lcccc}
             & Negative & Neutral & Positive \\\hline
    Training & 174 (46\%) & 162 (42\%) & 46 (12\%)\\
    Test     & 102 (33\%)& 182 (59\%) & 24 (8\%)
  \end{tabular}
  \caption{Manual annotation of training and test set.}
  \label{tab:1}
\end{table}

\subsection{Sentiment lexicon and sentiment shifters}

Unfortunately, no sentiment list exists for the Norwegian language and therefore we developed our own by manually translating the AFINN list \cite{Nielsen11}. We also manually added 1590 words relevant to political discussions  like 'deport', 'expel', 'extremist', 'terrorist' and so on, ending up with a list of 4067 Norwegian sentiment words. Each word were given a score from $-5$ to $5$ ranging from words with extremely negative sentiment (e.g. 'behead') to highly positive sentiment words (e.g. 'breathtaking').

Several Norwegian sentiment shifters were considered but only the basic shifter 'not' improved the opinion classification and therefore only this word were used in the method.

\subsection{classification performance}

In this study we compared four different cases based on the general algorithm in \eqref{eq:1}.
\begin{itemize}
\item We use the \textbf{imp}-function presented in \eqref{eq:2}. We denote this case WD (word distance).
\item For this case and the two below we use the \textbf{imp}-function in \eqref{eq:3}. Further we set $\alpha = 0$ meaning that we do not use the important dependence paths. We denote this case A0 ($\alpha = 0$).
\item We set $\alpha = 1$ and for all dependency paths we set $W_{\mathbf{dep}} = 2/3$. We denote this case CW (constant weights).
\item We set $\alpha = 1$ and for $W_{\mathbf{dep}}$ we use the weights presented in Table ??????. We denote this case OD (optimal use of dependence paths)
\end{itemize}
For each case we used the training set to manually tune the parameters $d_p, d_n, t_p$ and $t_n$ of the method. The parameters were tuned to optimize the number of correct classifications.

Table \ref{tab:2} shows the optimal parameter values of $d_p, d_n, t_p$ and $t_n$ and the classification performance for different cases. 
\begin{table}
  \centering
  \begin{tabular}{lcccccl}
       & $d_p$ & $d_n$ & $t_p$ & $t_n$ & Correct   & p-val \\ \hline
    WD &    2  &   0  & 0.7  &  0.0  & 142 (46\%) & \\
    A0 &    2  &   0  & 2.5  &  0.3  & 161 (52\%) & 0.014 *\\
    CW &    2  &   0  & 2.5  &  0.3  & 160 (52\%) & 0.020 *\\
    OD &    2  &   0  & 2.5  &  0.3  & 162 (53\%) & 0.009 **
  \end{tabular}
  \caption{The second to the fifth column show the optimal values of the parameters of the model. The sixth column show the number of correct classifications and the last column shows the p-value that the method performs better than WD. Signif. codes: p-value $<$ 0.01:\,**, \, p-value $<$ 0.05:\,*.}
  \label{tab:2}
\end{table}
We see that the sentiment shifter 'not' only have a positive effect on classification performance when it is in front of the sentiment word. We see that using dependence tree distances (case A0) the classification results is significantly improved compared word distance in the sentence (case WD) (p-value = 0.014). Also including the important dependence paths (case OD) further improves the classification performance, but the improvement is not significant compared to case A0.

@@

In this paper we have analyzed if parsing can improve
